Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes

### 2019-05-03 GWLDCI.tex
some little typos and putting admissible in the statement and now I'm happy with 3.17

\begin{dfn} \label{VPropdfn}
Let $W = (\cP, [n])$.
\begin{enumerate}
\item For $p \in \cP$, let $V(p) = \{i_p, i_p+1, j_p, j_p+1\}$ be the set of vertices supporting $p$. Then, for $P \subseteq \cP$, the set $V(P) = \cup_{p \in P} V(p)$ is the vertex support of $P$.
\item For $V \subseteq [n]$, write $\Prop(V) = \{ p \in \cP | V(p) \cap V \neq \emptyset \} $.
\item For $P \subseteq \cP$, define $F(P) = V(P^c)^c$ to be\footnote{Here, as throughout, we are using the subscript $c$ for the set theoretic complement.} the set of vertices in $[n]$ that do not support any propagators outside the set $P$.
\item The set of vertices that are not in the support of any propagotors is denoted  $F(\emptyset)$. Vertices in this set are called non-supporting.  
\end{enumerate}
\end{dfn}

@@

The previous definition makes sense, in fact, for any two polygon dissections of the same polygon, not just for those in the image of $\tau$..

@@

\begin{proof}
This follows from the definition of equivalent Wilson loop diagrams, Definition \ref{equivdfn}, the definition of retriangulations, Definition \ref{retriangulation}, and Lemma\ref{lem triang to exact}.
\end{proof}

@@

\begin{cor} \label{classifyflats}
  If $F$ is a flat of a Wilson loop diagram, it can be written as the disjoint union of a cyclic propagator flat and an independent set: \bas F =C \sqcup S \eas where $C$ the union of all circuits in S. \end{cor}

@@

\begin{lem} \label{maxexactcomplementrank}
Let $W = (\cP, [n])$ be a Wilson loop diagram, and $P \subseteq \cP$. Then: \begin{enumerate}
\item If $(P,V(P))$ is an exact subdiagram in $W$, then $F(P^c)$ is not an independent flat.
\item If $(P,V(P))$ is a maximal exact subdiagram in $W$, then $F(P^c)$ is a cyclic flat.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          \item If $(P,V(P))$ is an exact subdiagram in $W$, then $\rk(F(P^c)) = |P^c|$.
\end{enumerate}
\end{lem}

@@

Since $R \subset P$, $V(P) = V(R) \sqcup S$, where $S = V(P) \setminus V(R)$. Since $P$ and $R$ both define exact subdiagrams, we may write $|V(P)| = |V(R)| + |P \setminus R|$, and we have that $S$ is a vertex set of size $|P \setminus R|$. Furthermore, $F(R^c) = F(P^c) \sqcup S$. We show that $S$ is an independent set in the subdiagram $(P, V(P))$, and thus in $W$. Independence in $(P, V(P))$ implies that for any maximal independent set $\mathfrak{B} \subset F(P^c)$, the set $\mathfrak{B} \sqcup S$ is independent in $F(R^c)$. Therefore, \bas \rk (F(R^c)) \geq \rk (F(P^c)) + |S| = |P^c| + |P\setminus R| = |R^c|  \;.\eas By Lemma \ref{lem facts about WLD matroids}, $\rk (F(R^c)) \leq |R^c|$. Thus $\rk (F(R^c)) = |R^c|$ as desired.

It remains to show that $S$ is an independent set in the subdiagram $(P, V(P))$. Suppose not. Then, by Theorem \ref{thm WLD defines matroid} there is a subset $U \subset S$ such that $|\Prop(U)| < |U|$. Then the remaining propagators, $P \setminus \Prop(U)$ are supported entirely on $V(P) \setminus U$. That is \bas  V(P \setminus \Prop(U)) = V(P) \setminus U\; .\eas Comparing sizes of these sets, we get \bas |V(P \setminus \Prop(U))| = |V(P)| - |U| = |P| +3 -|U| > |P| +3 -|\Prop(U)| = |P \setminus \Prop(U)| +3\; ,\eas which violates admissibility. Therefore, $S$ is independent in $(P, V(P))$.

>>>>

\begin{dfn} \label{VPropdfn}
Let $W = (\cP, [n])$.
\begin{enumerate}
\item For $p \in \cP$, let $V(p) = \{i_p, i_p+1, j_p, j_p+1\}$ be the set of vertices supporting $p$. Then, for $P \subseteq \cP$, the set $V(P) = \cup_{p \in P} V(p)$ is the vertex support of $P$.
\item For $V \subseteq [n]$, write $\Prop(V) = \{ p \in \cP | V(p) \cap V \neq \emptyset \} $.
\item For $P \subseteq \cP$, define $F(P) = V(P^c)^c$ to be\footnote{Here, as throughout, we are using the superscript $c$ for the set theoretic complement.} the set of vertices in $[n]$ that do not support any propagators outside the set $P$.
\item The set of vertices that are not in the support of any propagotors is denoted  $F(\emptyset)$. Vertices in this set are called non-supporting.  
\end{enumerate}
\end{dfn}

@@

The previous definition makes sense, in fact, for any two polygon dissections of the same polygon, not just for those in the image of $\tau$.

@@

\begin{proof}
This follows from the definition of equivalent Wilson loop diagrams, Definition \ref{equivdfn}, the definition of retriangulations, Definition \ref{retriangulation}, and Lemma \ref{lem triang to exact}.
\end{proof}

@@

\begin{cor} \label{classifyflats}
  If $F$ is a flat of an admissible Wilson loop diagram, it can be written as the disjoint union of a cyclic propagator flat and an independent set: \bas F =C \sqcup S \eas where $C$ the union of all circuits in S. \end{cor}

@@

\begin{lem} \label{maxexactcomplementrank}
Let $W = (\cP, [n])$ be an admissible Wilson loop diagram, and $P \subseteq \cP$. Then: \begin{enumerate}
\item If $(P,V(P))$ is an exact subdiagram in $W$, then $F(P^c)$ is not an independent flat.
\item If $(P,V(P))$ is a maximal exact subdiagram in $W$, then $F(P^c)$ is a cyclic flat.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          \item If $(P,V(P))$ is an exact subdiagram in $W$, then $\rk(F(P^c)) = |P^c|$.
\end{enumerate}
\end{lem}

@@

Since $R \subset P$, $V(P) = V(R) \sqcup S$, where $S = V(P) \setminus V(R)$. Since $P$ and $R$ both define exact subdiagrams, we may write $|V(P)| = |V(R)| + |P \setminus R|$, and we have that $S$ is a vertex set of size $|P \setminus R|$. Furthermore, $F(R^c) = F(P^c) \sqcup S$. We will show that $S$ is an independent set in the subdiagram $(P, V(P))$, and thus in $W$. Supposing for the moment that $S$ is independent in $(P, V(P)$ this implies that for any maximal independent set $\mathfrak{B} \subset F(P^c)$, the set $\mathfrak{B} \sqcup S$ is independent in $F(R^c)$. Therefore, \bas \rk (F(R^c)) \geq |\mathcal{B}| + |S| = \rk (F(P^c)) + |S| = |P^c| + |P\setminus R| = |R^c|  \;.\eas By Lemma \ref{lem facts about WLD matroids}, $\rk (F(R^c)) \leq |R^c|$. Thus $\rk (F(R^c)) = |R^c|$ as desired.

It remains to show that $S$ is an independent set in the subdiagram $(P, V(P))$. Suppose not. Then, by Theorem \ref{thm WLD defines matroid} there is a subset $U \subseteq S$ such that $|\Prop(U)| < |U|$. Then the remaining propagators, $P \setminus \Prop(U)$ are supported entirely on $V(P) \setminus U$. That is \bas  V(P \setminus \Prop(U)) = V(P) \setminus U\; .\eas Comparing sizes of these sets, we get \bas |V(P \setminus \Prop(U))| = |V(P)| - |U| = |P| +3 -|U| > |P| +3 -|\Prop(U)| = |P \setminus \Prop(U)| +3\; ,\eas which violates admissibility. Therefore, $S$ is independent in $(P, V(P))$.
